A Study on Online Sequential Learning using Bandits

### 2017-12-11 Chapter6/intro.tex
Latest on 11/12/2017

    The rest of the chapter is organized as follows. We first state the notations, definitions, and assumptions required for this setting in section~\ref{psbandit:notations}. Then we define our problem statement in section~\ref{psbandit:probDef} and in section~\ref{psbandit:related} we discuss the related works in this setting. We elaborate our contributions in section~\ref{psbandit:contribution} and in section~\ref{psbandit:algorithm} we present the changepoint detection algorithms. Section~\ref{psbandit:expt} contains numerical simulations where we test our proposed algorithms and finally we summarize in section \ref{psbandit:conclusion}.

>>>>

    The rest of the chapter is organized as follows. We first state the notations, definitions, and assumptions required for this setting in section~\ref{psbandit:notations}. Then we define our problem statement in section~\ref{psbandit:probDef} and in section~\ref{psbandit:related} we discuss the related works in this setting. We elaborate our contributions in section~\ref{psbandit:contribution} and in section~\ref{psbandit:algorithm} we present the changepoint detection algorithms. Section~\ref{psbandit:results} contains our main result, section~\ref{psbandit:expt} contains numerical simulations where we test our proposed algorithms and finally we summarize in section \ref{psbandit:conclusion}.

### 2017-12-11 Chapter6/notations.tex
Latest on 8/12/2017
Latest on 11/12/2017

	The distribution associated with individual arm $i$ is denoted by $D_{i,\rho_{c_j:c_{j+1}}}$ for the segment $\rho_{c_j:c_{j+1}}$ whereas the reward drawn from that distribution for the $t$-th time instant is denoted by $X_{i,t}$. We assume all rewards are bounded in $[0,1]$. $n_{i,t_0:t_p}$ denotes the number of times arm $i$ has been pulled from $t_0$ to $t_p$ timesteps. $r_{i,t_{c_j:c_{j+1}}}$ denotes the expected mean of the reward distribution $D_{i,\rho_{c_j:c_{j+1}}}$ and $\hat{r}_{i,t_0:t_p}$ denotes the sample mean for the arm $i$ from $t_0$ to $t_p$ timesteps.

@@

\begin{discussion}
\label{dis:gap-delay}
Thus, assumption \ref{assm:space-gap} makes sure that all the changepoints $t_{c_j},\forall j=1,2,\ldots,G$ are sufficiently far apart to detect a change. A good learner tries to minimize the delay ($d$) in detection of a changepoint at $t_{c_j},\forall c_{j}\in\G$ by achieving atmost a delay of $d\left(t_{c_j} - t_{c_{j-1}}\right) \leq O\left( \dfrac{\log( t_{c_j} - t_{c_{j-1}} )}{\Delta_{i,c_j}^{2}}\right)$. Note, that when the gaps are same such that for all $i\in\A$, $\Delta_{i,c_j}^{chg}=\Delta_{\epsilon_0,c_j}^{chg}$ and $d\left(t_{c_j} - t_{c_{j-1}}\right) = \left( \dfrac{\log( t_{c_j} - t_{c_{j-1}} )}{\Delta_{\epsilon_0,c_j}^{2}}\right)$, then due to assumption \ref{assm:space-gap} we get,
\begin{align*}
\Delta_{\epsilon_0,c_j}^{chg}\geq \sqrt{\dfrac{1-\epsilon_0}{\epsilon_0}\dfrac{\log(t_{c_j} - t_{c_{j-1}})}{(t_{c_{j+1}} - t_{c_{j}})}}.
\end{align*}

@@

\begin{align*}
\Psi_{i,c_j} = \dfrac{\Delta^{opt}_{i,c_{j+1}}}{\Delta^{chg}_{i,c_j}}
\end{align*}

>>>>

	The distribution associated with individual arm $i$ is denoted by $D_{i,\rho_{c_j:c_{j+1}}}$ for the segment $\rho_{c_j:c_{j+1}}$ whereas the reward drawn from that distribution for the $t$-th time instant is denoted by $X_{i,t}$. We assume all rewards are bounded in $[0,1]$. $n_{i,t_0:t}$ denotes the number of times arm $i$ has been pulled from $t_0$ to $t$ timesteps. $r_{i,t_{c_j:c_{j+1}}}$ denotes the expected mean of the reward distribution $D_{i,\rho_{c_j:c_{j+1}}}$ and $\hat{r}_{i,t_0:t}$ denotes the sample mean for the arm $i$ from $t_0$ to $t$ timesteps.

@@

\begin{discussion}
\label{dis:gap-delay}
Thus, assumption \ref{assm:space-gap} makes sure that all the changepoints $t_{c_j},\forall j=1,2,\ldots,G$ are sufficiently far apart to detect a change. A good learner tries to minimize the delay ($d$) in detection of a changepoint at $t_{c_j},\forall c_{j}\in\G$ by achieving atmost a delay of $d\left(t_{c_j} - t_{c_{j-1}}\right) \leq O\left( \dfrac{\log( t_{c_j} - t_{c_{j-1}} )}{(\Delta^{chg}_{i,c_j})^{2}}\right)$. Note, that when the gaps are same such that for all $i\in\A$, $\Delta_{i,c_j}^{chg}=\Delta_{\epsilon_0,c_j}^{chg}$ and $d\left(t_{c_j} - t_{c_{j-1}}\right) = \left( \dfrac{\log( t_{c_j} - t_{c_{j-1}} )}{(\Delta^{chg}_{\epsilon_0,c_j})^{2}}\right)$, then due to assumption \ref{assm:space-gap} we get,
\begin{align*}
\Delta_{\epsilon_0,c_j}^{chg}\geq \sqrt{\left(\dfrac{1-\epsilon_0}{\epsilon_0}\right)\dfrac{\log(t_{c_j} - t_{c_{j-1}})}{(t_{c_{j+1}} - t_{c_{j}})}}.
\end{align*}

@@

\begin{align*}
\Psi_{i,c_j} = \dfrac{\Delta^{opt}_{i,c_{j}}}{\Delta^{chg}_{i,c_j}}
\end{align*}

### 2017-12-11 Chapter6/probDef.tex
Latest on 11/12/2017

Given the exploration-exploitation-changepoint dilemma, the objective of the learner in the piecewise-stochastic bandit problem is to minimize the cumulative regret until time $T$, which is defined as follows:
\begin{align*}
R_{T}=\sum_{t'=1}^T r_{i^*_{t'}} - \sum_{t'=1}^T r_{\mathbb{I}_t'}
%R_{t}=\sum_{t'=1}^t r_{i^*_{t'}} - \sum_{i=1}^K \left(r_{i}n_{i_{t'}\neq i^*_{t'},\forall t'=1:t}\right),
\end{align*}
where $T$ is the total number of rounds or horizon, $r_{i^*_{t'}}$ is the optimal arm at the $t'$ timestep and $r_{\mathbb{I}_{t'}}$ is the arm chosen by the learner at the $t'$ timestep. Let $n_{i_{t'}\neq i^*_{t'},\forall t'=1:T}$ is the number of times the learner has chosen arm $i$ up to round $T$ when it was not the optimal arm $i^*_{t'}$. The expected regret of an algorithm after $T$ rounds can be written as,

\begin{align*}
\E[R_{T}]&= \E\left[\sum_{t'=1}^T r_{i^*_{t'}} - \sum_{t'=1}^t r_{\mathbb{I}_t'}\right]\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&=\sum_{i = 1}^K\left(r_{i^*} - r_i\right) \E[n_{i_{t'}\neq i^*_{t'},\forall t'=1:T}] \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\overset{(a)}{=}\sum_{i = 1}^K\sum_{j=1}^{G}\Delta^{opt}_{i,c_j}\E[n_{i_{t',c_j}\neq i^*_{t',c_j},\forall t'=t_{c_{j-1}}:t_{c_j}}]
\end{align*}
where $(a)$ is obtained from assumption \ref{assm:chg-gap} and $\Delta^{opt}_{i,c_j}$ is defined as in Definition \ref{Def:opt-gap}.

>>>>

Given the exploration-exploitation-changepoint dilemma, the objective of the learner in the piecewise-stochastic bandit problem is to minimize the cumulative regret until time $T$, which is defined as follows:
\begin{align*}
R_{T}=\sum_{t'=1}^T r_{i^*_{t'}} - \sum_{t'=1}^T r_{\mathbb{I}_t'}
%R_{t}=\sum_{t'=1}^t r_{i^*_{t'}} - \sum_{i=1}^K \left(r_{i}n_{i_{t'}\neq i^*_{t'},\forall t'=1:t}\right),
\end{align*}
where $T$ is the total number of rounds or horizon, $r_{i^*_{t'}}$ is the optimal arm at the $t'$ timestep and $r_{\mathbb{I}_{t'}}$ is the arm chosen by the learner at the $t'$ timestep. Let $n_{i_{t'}\neq i^*_{t'},\forall t'=1:T}$ is the number of times the learner has chosen arm $i_{t'}$ up to round $T$ when it was not the optimal arm $i^*_{t'}$. The expected regret of an algorithm after $T$ rounds can be written as,

\begin{align*}
\E[R_{T}]&= \E\left[\sum_{t'=1}^T r_{i^*_{t'}} - \sum_{t'=1}^t r_{\mathbb{I}_t'}\right]\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&=\sum_{i = 1}^K\left(r_{i^*_{t'}} - r_{i_{t'}}\right) \E[n_{i_{t'}\neq i^*_{t'},\forall t'=1:T}] \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\overset{(a)}{=}\sum_{i = 1}^K\sum_{j=1}^{G}\Delta^{opt}_{i,c_j}\E[n_{i_{t',c_j}\neq i^*_{t',c_j},\forall t'=t_{c_{j-1}}:t_{c_j}}]
\end{align*}
where $(a)$ is obtained from Assumption \ref{assm:chg-gap} and $\Delta^{opt}_{i,c_j}$ is defined as in Definition \ref{Def:opt-gap}.
